\documentclass[a4paper,12pt,twoside]{book}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
 \usepackage{lipsum}% http://ctan.org/pkg/lipsum
 \usepackage{fancyhdr}% http://ctan.org/pkg/fancyhdr
 \fancyhf{}% Clear all headers/footers
 \renewcommand{\headrulewidth}{0pt}% No header rule
 \renewcommand{\footrulewidth}{0pt}% No footer rule
 \fancyhead[R]{{\em updated 7/17/2017 VF}}%
\pagestyle{empty}

\begin{document}
	\thispagestyle{fancy}%
\section*{RNA Quality Assessment: Bleach gel}
 \subsection*{Purpose:}
To assess quality of extracted RNAs, by examining for fragmented or shearing. Equivalent of a genomic DNA gel, except bleach helps degrade RNases in gel. Nanodrop quantification gives metrics of purity (260/280 and 260/230 ratios) and concentration, but not general quality. 
\\
\\Protocol modified from Aranda et al (2012) {\em Electrophoresis}
\subsection*{Reagents:}
In addition to regular gel requirements...
\begin{itemize}
	\item TAE 1X gel buffer solution
	\item Bleach (standard household)
\end{itemize}
 \subsection*{Instructions:}
\begin{enumerate}
	\item Mix 1 percent agarose with TAE 1X buffer. (e.g. 1 g agarose to 100 mL TAE buffer)
	\item To each 100 mL of gel volume, add 1000 uL of bleach using a pipette. 
	\item Swirl and heat gel mixture in the microwave as per usual protocol. 
	\item Add EtBR, pour,  set combs and let cool as per standard gel protocol. 
	\item To prepare RNA for loading, pipette out 6 uL of RNA onto a labeled sheet of parafilm. {em\ Keep RNA stocks cool during this step.} Add 1 uL 6X loading dye to each 6 uL RNA. Load 7 uL into gel combs. 
	\item Run for 100 Volts for 35 minutes. 
	\item {\bf Interpretation:}
	You should be able to see two (ideally, three) distict bands on your gel after running. The top band (~ 4000 bp) corresponds to 28S rRNA, while lower band (~2000 bp) corresponds to 18S rRNA. Much below, around 150 bp, is 5S rRNA. 5S rRNA tends to be more faint and "blurred" of a band. If all three bands are distnctly seen, then RNA is of high quality. Smudges and streaks suggest RNA degredation. Glowing wells suggest gDNA contamination.  
	
	
\end{enumerate}

\end{document}